\documentclass[12pt,letterpaper,twoside]{book}
\usepackage[utf8]{inputenc}
%\usepackage[width=150mm,top=25mm,bottom=25mm,bindingoffset=6mm]{geometry}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{mathtools}
\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{nameref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[nocfg,refpage]{nomencl}
\makenomenclature
\renewcommand{\nomname}{List of Symbols}
%\usepackage{biblatex}
\newcommand{\G}{G^F}
\newcommand{\supp}{\text{supp}}
\newcommand{\diam}{\text{diam}}
\newcommand{\Exc}{\text{Exc}}
\newcommand{\GO}{\text{GO}}
\newcommand{\C}{\widehat{\mathbb{C}}}
\newcommand{\D}{\mathbb{D}}
\renewcommand{\nompreamble}{In order of appearance.}
%Proposiciones
%\theoremstyle{plain}
\newtheorem{prop}{Proposition}[section]
\newtheorem{theo}{Theorem}[section]
%\newtheorem{defprop}[prop]{Definition-Proposition}
\newtheorem{lemma}{Lemma}[section]
\newtheorem{coro}{Corollary}[section]
%\theoremstyle{definition}
\newtheorem{mdef}{Definition}[section]
\theoremstyle{definition}
\newtheorem{exmp}{Example}[section]
\newtheorem{rmk}{Remark}[section]
\newtheorem{qstn}{Question}[section]
%Frontmatter
\usepackage[english]{babel} % language/hyphenation
%\usepackage[utf8]{inputenc}  % input encoding
\usepackage{graphicx} % graphics library (REQUIRED by CIMATpreamble)
\usepackage[svgnames,table]{xcolor}  % color library (REQUIRED by CIMATpreamble)
\usepackage{CIMATpreamble}  % definition of the CIMAT cover page
%\usepackage{mypreamble} % custom preamble
\hypersetup{
	colorlinks = true,
	linkcolor = blue,
	citecolor = red,
	}
\usepackage{tikz}
\usepackage[most]{tcolorbox}
\tcbuselibrary{theorems,breakable}
\newtcbtheorem[number within=section]{mytheo}{Theorem}%
{colback=pink!5,colframe=pink!35!black,fonttitle=\bfseries}{th}
\newtcbtheorem[number within=section]{myprop}{Proposition}%
{colback=pink!5,colframe=pink!35!black,fonttitle=\bfseries}{pr}
\newtcbtheorem[number within=section]{myexmp}{Example}%
{colback=pink!5,colframe=pink!35!black,fonttitle=\bfseries}{ex}
\newtcbtheorem[number within=section]{mylema}{Lemma}%
{colback=pink!5,colframe=pink!35!black,fonttitle=\bfseries}{lm}
\newtcbtheorem[number within=section]{mycoro}{Corollary}%
{colback=pink!5,colframe=pink!35!black,fonttitle=\bfseries}{cr}
\newtcbtheorem[number within=section]{myrmk}{Remark}%
{colback=pink!5,colframe=pink!35!black,fonttitle=\bfseries}{rm}
\newtcbtheorem[number within=section]{mydef}{Definition}%
{colback=pink!5,colframe=pink!35!black,fonttitle=\bfseries}{df}
\newtcbtheorem[number within=section]{myqstn}{Question}%
{colback=pink!5,colframe=pink!35!black,fonttitle=\bfseries}{qs}
\pagestyle{fancy}
\fancyhead{}
\fancyhead[RO]{\nouppercase {\rightmark}}
\fancyhead[LE]{\nouppercase {\leftmark}}
\fancyfoot{}
\fancyfoot[LE,RO]{\thepage}
\author{René Velázquez Ascencio}
\documentType{T E S I S}  % change to "T E S I N A" if it is the case
\title{On a measure theoretic characterization of polynomial Julia sets}
\degree{Maestro en Ciencias con Orientación en Matemáticas Básicas}
\supervisor{Dra. Mónica Moreno Rocha}
%\supervisorSecond{Dra. }  % optional command
\cityandyear{Guanajuato, Guanajuato, Diciembre de 2022}
\textheight = 22 cm
\textwidth = 17cm
\topmargin = -1cm
\oddsidemargin = .1cm
\evensidemargin = .3cm
\begin{document}
\maketitle


\chapter*{Introduction}

The Julia set $J(f)$ of a non-constant rational map $f$ of degree $d\geq 2$, acting on the Riemann sphere $\C$, can be considered as the set where the iterates of $f$ behave in a chaotic manner. We may ask if there are any conditions which could distinguish if a given Julia set came up from a polynomial map or from a genuine rational map, that is a meromorphic function on $\C$ with at least one finite pole. However, it is not sufficient to ask information only from $J(f)$, since there are polynomials and strict rational maps which have the same Julia set. For instance, if $f(z)=z^n(z-a)/(1-\overline{a}z)$, for $0<|a|<1$ and $n\geq 2$, then $J(f)$ is the unit circle, see Example \ref{ex:ejemploblashky}. But for any polynomial of the form $P(z)=z^n$ for $|n|>1$, the Julia set $J(P)$ is also the unit circle, see Example \ref{ex:zalan}. Nonetheless, when considering not only the Julia set but also dynamical information generated by the rational map, some interesting results have been found.\\

In 1965 Brolin \cite{brolin1965invariant} proved that, if $f$ is a polynomial, then the equilibrium measure $\nu$ associated to $J(f)$ and the measure  of maximal entropy $\mu_f$ of $f$ are equal. Intuitively, the equilibrium measure of a compact set $K\subset \mathbb{C}$ resembles a mass charge distribution produced when a set of charges reaches an equilibrium in $K$. Formally, the logarithmic potential with respect to the equilibrium measure is constant {\it nearly} everywhere in $K$, see Definition \ref{df:nearlyeverywhere} and Theorem \ref{th:frostman}. On the other hand the measure of maximal entropy $\mu_f$ measures the proportion of preimages in $f^{-n}(a)$ that accumulate about a given subset of $J(f)$, when $n\rightarrow \infty$ and $a$ has an infinite backward orbit. So the problem stated above would have an answer if there were a converse to Brolin's Theorem. It can be shown that, strictly speaking, the converse to Brolin's Theorem is false, see Remark \ref{rm:conversetobrolin}. However, if $f$ were a polynomial then $\infty$ would be an attracting fixed point and this implies that $\infty\not \in J(f)$. Hence, before trying to check if a rational map $f$ is a polynomial, a natural condition to impose is $f(\infty)=\infty$, since otherwise the answer would be trivially negative. Indeed, in 1984 Lopes \cite{lopes} proved that for a rational map $f$ satisfying $f(\infty)=\infty\not \in J(f)$, if the equilibrium measure and the maximum entropy measure coincide then $f$ is a polynomial. In 2011 Okuyama and Stawiska \cite{okuyama} relaxed Lopes' hypothesis by requiring that $\infty$ lies in a forward invariant Fatou component, see Definitions \ref{df:JuliaF} and \ref{df:definvariance}. Moreover, in 2018 they weakened their own hypothesis a bit more by showing that if $J(f)$ is bounded, then $\nu=\mu_f$ if and only if $f^2$ is a polynomial \cite{okuyama2018}. Let us state now the main theorem that we shall study in this work.

\begin{theo}[ Okuyama \& Stawiska]\label{theo:Oku}
Let $f:\C \rightarrow \C$ be a rational map of degree $d\geq 2$. If $\infty$ lies in a forward invariant Fatou component and $\mu_f=\nu$ then $f$ is a polynomial. Conversely, if $f$ is a polynomial then $\mu_f=\nu$
\end{theo}

The present work is aimed to give enough details to comprehend Theorem \ref{theo:Oku}. The structure of the thesis is divided as follows. In Chapter \ref{chapter1} we collect some necessary theory to understand the methods used in the article by Okuyama and Stawiska. It covers basic complex dynamics, analysis in several complex variables, potental theory, etc.. Then, in Chapter \ref{chapter2} we prove the first statement of Theorem \ref{theo:Oku}. In Chapter \ref{chapter3} we prove the converse direction, which simply states that if $f$ is a polynomial then $\mu_f=\nu$. This chapter also includes a discussion of Brolin's proof. Finally, in Chapter \ref{chapter 4} we explore several consequences of the theory covered in Chapters \ref{chapter2} and \ref{chapter3}. For example, given a rationa map $f$, using Lemma \ref{lm:pre_lema3} we were able to find an upper bound for the logarithmic potential $p_{\mu_f}$ restricted to the Julia set. Particularly, when $f$ is a polynomial, studying this upper bound guided us to find another proof of the equilavence $\mu_f=\nu$. The proof relies strongly in Lemma \ref{lm:pre_lema1}, proved in \cite{okuyama}. Chapter \ref{chapter 4} also includes several queries or questions that we found interesting for further exploration on the subject. In the appendices to this work is collected complementary theory which comprehends a brief discussion of Montel's Theorem, the proof of existence of lifts for rational functions, analytic curves and ultimately the theory of currents, a technical device from complex analysis in several variables.

\chapter*{Acknowledgments}
I am very thankful to many people who gave me their support to complete this work.\\

First of all I must thank my advisor Dra. Mónica Moreno for accepting me as one of her students even when by bachelor's background was not focused on dynamical systems. It has been a very stisfactory experience working with her. I am very grateful for her support as a professor and friend. She took a lot of time teaching and discussing ideas with me, not to say assiting me in redacting this work.\\

There was a moment when I was stuck trying to understand a certain proof, so I dared to email professor Margaret Stawiska. She was incredibly kind and gave me a lot of assistance. She suggested me additional bibliographic resources so I could understand the identity theorem of analytical paths. She did also provide me with hints and explanantions with which I was able to write Appendix \ref{sect3}. Very generously, she accepted to be a special reader in the thesis committee of this work and wrote a report with very important observations to improve this dissertation. I am also thankful to my sinodals Dr. Xavier Gómez-Mont and Dr. Renato Iturriaga, for reviewing this thesis, for their comments and for asking very interesting questions. I would like to thank especially Dr. Xavier for having me as a listener in his 2021 summer course on Potential Theory, those classes were very useful to me. He did also provide me with advices and support during the rest of my stay at CIMAT, for which I am very grateful.\\

I have to thank my family and friends as well. Thanks to my mom for her unconditional love and support which has been fundamental my whole life. Thanks to my old friend Jair for being my roomie and all the fun we had (toro lml), and for enduring my presentations. Thanks to Alejandra who helped me a lot when I arrived to Guanajuato, not every day someone does as much as she did for me on those days. Thanks to all the guys who gave me moments of fun in Guanajuato. Thanks to everyone who supported me but I'm not explicitly mentioning here.\\

Thanks to the National Council of Science and Technology (CONACYT) for the scholarship that funded my expenses and allowed me to be fully dedicated to obtain my master's degree, this was indeed fundamental. I hope the mathematical community in México will continue growing with the aid of this kind of support. Finally, thanks to Dr. Manuel González Villa, for including me in his research project, so I could complete this work.


\tableofcontents

\printnomenclature

\chapter{Introductory theory}\label{chapter1}

\input{Chapters/Chapter1}

\chapter{If $\mu_f=\nu$ and $f(D_\infty)\subset D_\infty$ then $f$ is a polynomial  }\label{chapter2}

\input{Chapters/Chapter2}

\chapter{If $f$ is a polynomial then $\mu_f=\nu$}\label{chapter3}

\input{Chapters/Chapter3}


\chapter{Our contributions and questions}\label{chapter 4}
\input{Chapters/conclusions}

\appendix
\chapter{Appendix}\label{chapterappendix}

\input{Chapters/appendix}

\bibliographystyle{siam}
\bibliography{referencias}
\end{document}
